---
title: "PNAD Contínua — Survey Analysis in R"
subtitle: "Descriptive Analysis, Survey Weights, and Regression Models"
author: "Teresa De Bastiani"
date: "`r format(Sys.Date(), '%B %d, %Y')`"
output:
  html_document:
    theme: flatly
    toc: true
    toc_float: true
    toc_depth: 3
    code_folding: show
    fig_width: 9
    fig_height: 5
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo    = TRUE,
  message = FALSE,
  warning = FALSE,
  comment = ""
)
```

---

## Overview

This document presents a three-part analysis of Brazil's national household survey
(**PNAD Contínua**, 4th quarter 2023) using microdata downloaded directly from IBGE
via the `PNADcIBGE` package.

The analysis follows a methodological progression:

1. **Descriptive analysis** — distributions, cross-tabulations, and visualization
2. **Survey weights** — why unweighted estimates are biased and how to correct them
3. **Regression models** — OLS and logistic models with correct inference for complex survey data

All categorical variables use the labels returned by `get_pnadc(labels = TRUE)`,
which applies the official IBGE data dictionary. No numeric codes are used.

---

## 1. Setup

```{r libraries}
library(PNADcIBGE)
library(tidyverse)
library(survey)
library(srvyr)
library(broom)

options(survey.lonely.psu = "adjust")

theme_pnadc <- function() {
  theme_minimal(base_size = 12) +
    theme(
      plot.title       = element_text(face = "bold", size = 13),
      plot.subtitle    = element_text(color = "grey40", size = 10),
      plot.caption     = element_text(color = "grey55", size = 8),
      axis.title       = element_text(size = 10),
      legend.position  = "bottom",
      panel.grid.minor = element_blank()
    )
}

PALETTE <- c(
  "main"   = "#2E86AB",
  "accent" = "#E24B4A",
  "mid"    = "#F18F01",
  "soft"   = "#B0C4DE",
  "green"  = "#4CAF50"
)
```

---

## 2. Data download

PNAD Contínua microdata is downloaded directly from IBGE servers using `get_pnadc()`.
Setting `labels = TRUE` applies the official data dictionary to all categorical variables,
returning them as labeled factors. This eliminates the need for manual numeric recoding
and makes the code robust to dictionary changes across survey editions.

```{r download, cache=TRUE}
vars_needed <- c(
  "UF", "UPA", "Estrato", "V1028",
  "V2007", "V2009", "V2010",
  "VD3005", "VD4001", "VD4002", "VD4009", "VD4020"
)

pnadc_design <- get_pnadc(
  year     = 2023,
  quarter  = 4,
  vars     = vars_needed,
  labels   = TRUE,
  deflator = FALSE,
  design   = TRUE
)

cat(sprintf("Download complete: %d respondents\n", nrow(pnadc_design$variables)))
```

---

## 3. Label inspection

Before recoding any variable, we inspect the labels returned by the IBGE dictionary.
The exact strings shown here are what must be used in all `case_when()` calls below.

```{r labels}
cat("V2007 (Gender):\n");      print(levels(pnadc_design$variables$V2007))
cat("V2010 (Race):\n");        print(levels(pnadc_design$variables$V2010))
cat("VD3005 (Schooling):\n");  print(levels(pnadc_design$variables$VD3005))
cat("VD4001 (Labor force):\n"); print(levels(pnadc_design$variables$VD4001))
cat("VD4002 (Employment):\n"); print(levels(pnadc_design$variables$VD4002))
cat("VD4009 (Job type):\n");   print(levels(pnadc_design$variables$VD4009))
```

---

## 4. Variable recoding

All recoding is done inside the design object using `update()`. Extracting the data,
recoding, and rebuilding the design would break the sample design structure.

```{r recoding}
pnadc_design <- update(pnadc_design,

  education = factor(case_when(
    VD3005 == "Sem instrução e menos de 1 ano de estudo" ~ "No schooling",
    VD3005 %in% c(
      "1 ano de estudo","2 anos de estudo","3 anos de estudo","4 anos de estudo"
    ) ~ "Primary incomplete",
    VD3005 %in% c(
      "5 anos de estudo","6 anos de estudo","7 anos de estudo","8 anos de estudo"
    ) ~ "Primary complete",
    VD3005 %in% c(
      "9 anos de estudo","10 anos de estudo","11 anos de estudo"
    ) ~ "Secondary",
    VD3005 %in% c(
      "12 anos de estudo","13 anos de estudo","14 anos de estudo",
      "15 anos de estudo","16 anos ou mais de estudo"
    ) ~ "Higher education",
    TRUE ~ NA_character_
  ), levels = c("No schooling","Primary incomplete","Primary complete",
                "Secondary","Higher education")),

  race = factor(case_when(
    V2010 == "Branca"   ~ "White",
    V2010 == "Preta"    ~ "Black",
    V2010 == "Parda"    ~ "Brown (Pardo)",
    V2010 == "Amarela"  ~ "Asian",
    V2010 == "Indígena" ~ "Indigenous",
    TRUE                ~ NA_character_
  ), levels = c("White","Black","Brown (Pardo)","Asian","Indigenous")),

  macro_region = factor(case_when(
    UF %in% c("Rondônia","Acre","Amazonas","Roraima","Pará","Amapá","Tocantins") ~ "North",
    UF %in% c("Maranhão","Piauí","Ceará","Rio Grande do Norte","Paraíba",
              "Pernambuco","Alagoas","Sergipe","Bahia") ~ "Northeast",
    UF %in% c("Minas Gerais","Espírito Santo","Rio de Janeiro","São Paulo") ~ "Southeast",
    UF %in% c("Paraná","Santa Catarina","Rio Grande do Sul") ~ "South",
    UF %in% c("Mato Grosso do Sul","Mato Grosso","Goiás","Distrito Federal") ~ "Center-West",
    TRUE ~ NA_character_
  ), levels = c("Southeast","North","Northeast","South","Center-West")),

  age    = as.numeric(V2009),
  age_sq = as.numeric(V2009)^2,

  log_income = ifelse(!is.na(VD4020) & VD4020 > 0, log(VD4020), NA_real_),

  in_pea = !is.na(VD4001) & VD4001 == "Pessoas na força de trabalho",

  employed = case_when(
    VD4002 == "Pessoas ocupadas"    ~ 1L,
    VD4002 == "Pessoas desocupadas" ~ 0L,
    TRUE ~ NA_integer_
  )
)

employed_design <- subset(pnadc_design, employed == 1L & !is.na(log_income))
pea_design      <- subset(pnadc_design, in_pea == TRUE & !is.na(employed))

# Raw data for descriptive plots
pnadc <- pnadc_design$variables

cat(sprintf("Employed with income: %d\nPEA: %d\n",
            nrow(employed_design$variables),
            nrow(pea_design$variables)))
```

---

## 5. Descriptive analysis

### 5.1 Income distribution by gender

The distribution of log monthly income for employed workers shows a clear rightward
shift for men — reflecting a persistent gender income gap across the entire distribution,
not just at the mean.

```{r plot-income-gender, fig.cap="Income distribution by gender (log scale)"}
pnadc |>
  filter(employed == 1L, !is.na(log_income), !is.na(V2007)) |>
  mutate(V2007 = as.character(V2007)) |>
  ggplot(aes(x = log_income, fill = V2007)) +
  geom_density(alpha = 0.55, color = NA) +
  scale_fill_manual(values = c("Homem" = unname(PALETTE["main"]),
                               "Mulher" = unname(PALETTE["accent"]))) +
  labs(
    title   = "Income distribution by gender",
    subtitle = "Employed population — log scale | PNAD Contínua 4Q2023",
    x = "Log of monthly income (R$)", y = "Density", fill = NULL,
    caption = "Source: IBGE, PNAD Contínua. Author's own analysis."
  ) +
  theme_pnadc()
```

### 5.2 Mean income by education level

Income increases sharply with education. Workers with higher education earn more than
twice the mean income of those with primary complete education.

```{r plot-income-education, fig.cap="Mean income by education level"}
pnadc |>
  filter(employed == 1L, !is.na(VD4020), VD4020 > 0, !is.na(education)) |>
  group_by(education) |>
  summarise(mean_income = mean(VD4020), .groups = "drop") |>
  ggplot(aes(x = education, y = mean_income, fill = education)) +
  geom_col(width = 0.65, show.legend = FALSE) +
  geom_text(aes(label = scales::dollar(mean_income, prefix = "R$",
                                        big.mark = ".", decimal.mark = ",")),
            vjust = -0.4, size = 3.2) +
  scale_fill_manual(values = colorRampPalette(c(PALETTE["soft"], PALETTE["main"]))(5)) +
  scale_y_continuous(
    labels = scales::dollar_format(prefix = "R$", big.mark = ".", decimal.mark = ","),
    expand = expansion(mult = c(0, 0.12))
  ) +
  labs(
    title    = "Mean monthly income by education level",
    subtitle = "Employed population | PNAD Contínua 4Q2023",
    x = NULL, y = "Mean income (R$)",
    caption  = "Source: IBGE, PNAD Contínua. Author's own analysis."
  ) +
  theme_pnadc() +
  theme(axis.text.x = element_text(angle = 20, hjust = 1))
```

### 5.3 Mean income by race and gender

The racial income gap is substantial and consistent across genders. White workers
earn more than Black and Brown workers in every gender group.

```{r plot-income-race, fig.cap="Mean income by race and gender"}
pnadc |>
  filter(employed == 1L, !is.na(VD4020), VD4020 > 0,
         !is.na(V2010), !is.na(V2007),
         V2010 %in% c("Branca","Preta","Parda")) |>
  group_by(V2010, V2007) |>
  summarise(mean_income = mean(VD4020), .groups = "drop") |>
  mutate(V2007 = as.character(V2007), V2010 = as.character(V2010)) |>
  ggplot(aes(x = reorder(V2010, mean_income), y = mean_income, fill = V2007)) +
  geom_col(position = "dodge", width = 0.65) +
  scale_fill_manual(values = c("Homem" = unname(PALETTE["main"]),
                               "Mulher" = unname(PALETTE["accent"]))) +
  scale_y_continuous(
    labels = scales::dollar_format(prefix = "R$", big.mark = ".", decimal.mark = ","),
    expand = expansion(mult = c(0, 0.12))
  ) +
  coord_flip() +
  labs(
    title    = "Mean income by race and gender",
    subtitle = "Employed population | PNAD Contínua 4Q2023",
    x = NULL, y = "Mean monthly income (R$)", fill = NULL,
    caption  = "Source: IBGE, PNAD Contínua. Author's own analysis."
  ) +
  theme_pnadc()
```

### 5.4 Mean income by macro-region

Regional income inequality is pronounced. Northeast workers earn on average 46% less
than workers in the South and Southeast.

```{r plot-income-region, fig.cap="Mean income by macro-region with 95% CI"}
region_summary <- pnadc |>
  filter(employed == 1L, !is.na(VD4020), VD4020 > 0, !is.na(macro_region)) |>
  group_by(macro_region) |>
  summarise(
    mean_income = mean(VD4020),
    se          = sd(VD4020) / sqrt(n()),
    ci_lower    = mean_income - 1.96 * se,
    ci_upper    = mean_income + 1.96 * se,
    .groups     = "drop"
  )

region_summary |>
  ggplot(aes(x = reorder(macro_region, mean_income), y = mean_income,
             fill = reorder(macro_region, mean_income))) +
  geom_col(width = 0.65, alpha = 0.9, show.legend = FALSE) +
  geom_errorbar(aes(ymin = ci_lower, ymax = ci_upper),
                width = 0.25, color = "grey30", linewidth = 0.7) +
  geom_text(aes(label = scales::dollar(mean_income, prefix = "R$",
                                        big.mark = ".", decimal.mark = ",")),
            hjust = -0.15, size = 3.2, color = "grey20") +
  scale_fill_manual(values = colorRampPalette(c(PALETTE["soft"], PALETTE["main"]))(5)) +
  scale_y_continuous(
    labels = scales::dollar_format(prefix = "R$", big.mark = ".", decimal.mark = ","),
    expand = expansion(mult = c(0, 0.20))
  ) +
  coord_flip() +
  labs(
    title    = "Mean monthly income by macro-region",
    subtitle = "Employed population | Error bars = 95% CI | PNAD Contínua 4Q2023",
    x = NULL, y = "Mean monthly income (R$)",
    caption  = "Source: IBGE, PNAD Contínua. Author's own analysis."
  ) +
  theme_pnadc() +
  theme(panel.grid.major.y = element_blank())
```

---

## 6. Survey weights and complex sample design

PNAD Contínua uses a complex probability sample — it is not a simple random sample.
Ignoring the sample design leads to biased estimates and underestimated standard errors.

Every estimate must account for:

- **Sampling weight** (`V1028`): how many people each respondent represents
- **Stratum** (`Estrato`): groups within which sampling was independent
- **PSU** (`UPA`): primary sampling unit (cluster)

### 6.1 Weighted vs. unweighted comparison

```{r weighted-comparison}
fmt_brl <- function(x) format(round(x, 2), big.mark = ".", decimal.mark = ",", nsmall = 2)

naive_mean   <- mean(employed_design$variables$VD4020, na.rm = TRUE)
weighted_est <- svymean(~VD4020, employed_design, na.rm = TRUE)
ci           <- confint(weighted_est)

cat(sprintf("  Unweighted mean: R$ %s\n", fmt_brl(naive_mean)))
cat(sprintf("  Weighted mean:   R$ %s\n", fmt_brl(coef(weighted_est))))
cat(sprintf("  SE (weighted):   R$ %s\n", fmt_brl(SE(weighted_est))))
cat(sprintf("  95%% CI: R$ %s to R$ %s\n", fmt_brl(ci[1]), fmt_brl(ci[2])))
```

The weighted mean (R$ 3,063) is 11% higher than the unweighted mean (R$ 2,763).
This difference arises because the sample overrepresents lower-income strata.
Reporting unweighted estimates for a nationally representative survey is methodologically incorrect.

### 6.2 Weighted mean income by education

```{r weighted-education}
svyby(~VD4020, ~education, employed_design, svymean, na.rm = TRUE) |>
  as_tibble() |>
  set_names(c("education", "mean_income", "se")) |>
  mutate(
    ci_lower = mean_income - 1.96 * se,
    ci_upper = mean_income + 1.96 * se,
    cv_pct   = round(se / mean_income * 100, 1)
  ) |>
  arrange(desc(mean_income))
```

```{r plot-weighted-education, fig.cap="Weighted mean income by education with 95% CI"}
income_edu <- svyby(~VD4020, ~education, employed_design, svymean, na.rm = TRUE) |>
  as_tibble() |>
  set_names(c("education", "mean_income", "se")) |>
  mutate(
    ci_lower = mean_income - 1.96 * se,
    ci_upper = mean_income + 1.96 * se
  ) |>
  filter(!is.na(education))

ggplot(income_edu,
       aes(x = factor(education,
                      levels = c("No schooling","Primary incomplete",
                                 "Primary complete","Secondary","Higher education")),
           y = mean_income)) +
  geom_col(fill = unname(PALETTE["main"]), width = 0.65, alpha = 0.9) +
  geom_errorbar(aes(ymin = ci_lower, ymax = ci_upper),
                width = 0.2, color = unname(PALETTE["accent"]), linewidth = 0.8) +
  scale_y_continuous(
    labels = scales::dollar_format(prefix = "R$", big.mark = ".", decimal.mark = ","),
    expand = expansion(mult = c(0, 0.12))
  ) +
  labs(
    title    = "Weighted mean income by education level",
    subtitle = "Error bars = 95% confidence interval | PNAD Contínua 4Q2023",
    x = NULL, y = "Weighted mean monthly income (R$)",
    caption  = "Source: IBGE, PNAD Contínua. Complex sample design incorporated."
  ) +
  theme_pnadc() +
  theme(axis.text.x = element_text(angle = 20, hjust = 1))
```

### 6.3 Design effect (DEFF)

The design effect measures how much the complex sample inflates variance compared to
a simple random sample of the same size. DEFF = 4.90 in the North means the effective
sample size for North regional estimates is less than a quarter of the nominal count.

```{r deff}
deff_df <- svyby(~VD4020, ~macro_region, employed_design,
                 svymean, na.rm = TRUE, deff = TRUE) |>
  as_tibble() |>
  set_names(c("macro_region", "mean_income", "se", "deff")) |>
  mutate(cv_pct = round(se / mean_income * 100, 1)) |>
  select(macro_region, mean_income, se, cv_pct, deff)

deff_df
```

```{r plot-deff, fig.cap="Design effect by macro-region", fig.height=4}
deff_df |>
  ggplot(aes(x = reorder(macro_region, deff), y = deff)) +
  geom_col(fill = unname(PALETTE["mid"]), width = 0.65, alpha = 0.9) +
  geom_hline(yintercept = 1, linetype = "dashed", color = "grey50") +
  geom_text(aes(label = round(deff, 2)), hjust = -0.2, size = 3.5) +
  coord_flip() +
  labs(
    title    = "Design effect (DEFF) by macro-region",
    subtitle = "DEFF > 1 = variance inflation vs. simple random sampling",
    x = NULL, y = "Design effect",
    caption  = "Source: IBGE, PNAD Contínua 4Q2023. Author's own analysis."
  ) +
  theme_pnadc()
```

### 6.4 Weighted cross-tabulation: employment by gender

```{r crosstab-gender}
as_survey_rep(pnadc_design) |>
  filter(!is.na(employed), !is.na(V2007)) |>
  group_by(V2007, as.character(employed)) |>
  summarise(pct = survey_mean() * 100, .groups = "drop") |>
  mutate(across(where(is.numeric), \(x) round(x, 1)))
```

---

## 7. Regression models

Both models use `svyglm()`, which correctly incorporates the complex sample design
into the estimation — producing valid standard errors, confidence intervals, and
p-values for population-level inference.

### 7.1 OLS — determinants of log income

**Model:** `log(income) ~ education + gender + race + age + age² + macro_region`

Coefficients on the log scale can be interpreted as approximate percentage changes
in income: `(exp(coef) - 1) × 100`.

Reference categories: Secondary (education), Homem (gender), White (race), Southeast (region).

```{r ols-model}
# Relevel for regression reference categories
pnadc_design <- update(pnadc_design,
  education    = relevel(education, ref = "Secondary"),
  race         = relevel(race,      ref = "White"),
  macro_region = relevel(macro_region, ref = "Southeast")
)
employed_design <- subset(pnadc_design, employed == 1L & !is.na(log_income))

model_ols <- svyglm(
  log_income ~ education + V2007 + race + age + age_sq + macro_region,
  design = employed_design,
  family = gaussian()
)

tidy_ols <- tidy(model_ols, conf.int = TRUE) |>
  filter(term != "(Intercept)") |>
  mutate(
    pct_effect = round((exp(estimate) - 1) * 100, 2),
    sig = case_when(
      p.value < 0.001 ~ "***",
      p.value < 0.01  ~ "**",
      p.value < 0.05  ~ "*",
      p.value < 0.1   ~ ".",
      TRUE             ~ ""
    )
  )

tidy_ols |>
  select(term, estimate, std.error, pct_effect, p.value, sig) |>
  mutate(across(where(is.numeric), \(x) round(x, 4)))
```

```{r plot-ols, fig.cap="OLS coefficient plot", fig.height=7}
tidy_ols |>
  mutate(
    direction  = ifelse(estimate > 0, "positive", "negative"),
    term_clean = term |>
      str_remove("^education|^V2007|^race|^macro_region") |>
      str_replace("age_sq", "Age²") |>
      str_replace("^age$", "Age")
  ) |>
  ggplot(aes(x = estimate, y = reorder(term_clean, estimate), color = direction)) +
  geom_vline(xintercept = 0, linetype = "dashed", color = "grey50") +
  geom_point(size = 3) +
  geom_errorbar(aes(xmin = conf.low, xmax = conf.high),
                width = 0.3, orientation = "y") +
  scale_color_manual(
    values = c("positive" = unname(PALETTE["main"]),
               "negative" = unname(PALETTE["accent"])),
    guide  = "none"
  ) +
  labs(
    title    = "OLS regression — determinants of log income",
    subtitle = "Coefficients with 95% CI | svyglm() with PNAD Contínua design",
    x = "Coefficient (log income scale)", y = NULL,
    caption  = "Reference: Secondary (education), Homem (gender), White (race), Southeast (region).\nSource: IBGE, PNAD Contínua 4Q2023."
  ) +
  theme_pnadc()
```

**Key findings:**

- Workers with **higher education** earn **70% more** than those with secondary education, controlling for all other factors
- **Women** earn **28% less** than men with equivalent education, race, age, and region
- **Black** and **Brown (Pardo)** workers earn approximately **20% less** than White workers
- **Northeast** workers earn **38% less** than Southeast workers

### 7.2 Logistic — probability of employment in the PEA

**Model:** `P(employed = 1) ~ education + gender + race + age + age² + macro_region`

The model is restricted to the **economically active population** (PEA) — people who are
either employed or actively seeking employment. This is the correct population for modeling
employment probability. Including inactive people (retired, students, homemakers) would
dilute the relationship between predictors and employment status.

Results are reported as **odds ratios** (OR > 1 = higher odds of employment).

```{r logit-model}
pea_design <- subset(pnadc_design, in_pea == TRUE & !is.na(employed))

model_logit <- svyglm(
  employed ~ education + V2007 + race + age + age_sq + macro_region,
  design = pea_design,
  family = quasibinomial()
)

tidy_logit <- tidy(model_logit, conf.int = TRUE, exponentiate = TRUE) |>
  filter(term != "(Intercept)") |>
  mutate(
    sig = case_when(
      p.value < 0.001 ~ "***",
      p.value < 0.01  ~ "**",
      p.value < 0.05  ~ "*",
      p.value < 0.1   ~ ".",
      TRUE             ~ ""
    )
  )

tidy_logit |>
  rename(odds_ratio = estimate, or_ci_low = conf.low, or_ci_high = conf.high) |>
  mutate(across(where(is.numeric), \(x) round(x, 4)))
```

```{r plot-logit, fig.cap="Logistic regression — odds ratios for employment", fig.height=7}
tidy_logit |>
  mutate(
    direction  = ifelse(estimate > 1, "higher odds", "lower odds"),
    term_clean = term |>
      str_remove("^education|^V2007|^race|^macro_region") |>
      str_replace("age_sq", "Age²") |>
      str_replace("^age$", "Age")
  ) |>
  ggplot(aes(x = estimate, y = reorder(term_clean, estimate), color = direction)) +
  geom_vline(xintercept = 1, linetype = "dashed", color = "grey50") +
  geom_point(size = 3) +
  geom_errorbar(aes(xmin = conf.low, xmax = conf.high),
                width = 0.3, orientation = "y") +
  scale_x_log10() +
  scale_color_manual(
    values = c("higher odds" = unname(PALETTE["green"]),
               "lower odds"  = unname(PALETTE["accent"])),
    guide  = "none"
  ) +
  labs(
    title    = "Logistic regression — odds ratios for employment",
    subtitle = "OR with 95% CI (log scale) | svyglm() with PNAD Contínua design",
    x = "Odds ratio (log scale) — OR > 1 = higher probability of employment",
    y = NULL,
    caption  = "Reference: Secondary (education), Homem (gender), White (race), Southeast (region).\nSource: IBGE, PNAD Contínua 4Q2023."
  ) +
  theme_pnadc()
```

```{r plot-edu-or, fig.cap="Employment odds ratios by education level", fig.height=4}
tidy_logit |>
  filter(str_detect(term, "^education")) |>
  mutate(
    term_clean = str_remove(term, "^education"),
    term_clean = factor(term_clean,
      levels = c("No schooling","Primary incomplete","Primary complete","Higher education")),
    direction  = ifelse(estimate > 1, "higher odds", "lower odds")
  ) |>
  ggplot(aes(x = estimate, y = reorder(term_clean, estimate), color = direction)) +
  geom_vline(xintercept = 1, linetype = "dashed", color = "grey50") +
  geom_point(size = 4) +
  geom_errorbar(aes(xmin = conf.low, xmax = conf.high),
                width = 0.25, orientation = "y", linewidth = 0.8) +
  geom_text(aes(label = sprintf("OR = %.2f", estimate)),
            hjust = -0.2, size = 3.2, color = "grey20") +
  scale_x_log10(limits = c(0.5, 2.5)) +
  scale_color_manual(
    values = c("higher odds" = unname(PALETTE["main"]),
               "lower odds"  = unname(PALETTE["accent"])),
    guide  = "none"
  ) +
  labs(
    title    = "Employment odds ratios by education level",
    subtitle = "Reference category: Secondary | PEA | PNAD Contínua 4Q2023",
    x = "Odds ratio (log scale) — OR > 1 = higher odds of employment vs. Secondary",
    y = NULL,
    caption  = "Source: IBGE, PNAD Contínua. svyglm() with quasibinomial() and complex sample design."
  ) +
  theme_pnadc() +
  theme(panel.grid.major.y = element_blank())
```

**Key findings:**

- **Women** in the labor force have **43% lower odds** of being employed than men (OR = 0.57), the strongest predictor in the model
- Workers with **higher education** have **57% higher odds** of employment than those with secondary education (OR = 1.57)
- **Black** workers have 22% lower odds of employment than White workers (OR = 0.78)
- The **South** shows substantially higher employment odds than the Southeast reference (OR = 1.68)

---

## 8. Summary

This analysis demonstrates a complete survey data workflow using PNAD Contínua:

| Pipeline | Key demonstration |
|---|---|
| Descriptive | Data download, label-based recoding, ggplot2 visualization |
| Survey weights | Correct weighted estimation, DEFF, `svyby()`, `srvyr` |
| Regression | `svyglm()` for OLS and logistic models with valid inference |

The methodological core of this project is the comparison between naive unweighted
analysis and design-corrected estimates. With design effects reaching 4.90 in the North,
ignoring the sample structure would produce confidence intervals that are less than half
the correct width for regional estimates.

---

*Data: IBGE, PNAD Contínua 4Q2023. Analysis: Teresa De Bastiani.*
